---
title: "DE"
author: "Nadine Bestard"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(Seurat) 
library(MAST)
library(tidyseurat)
library(dplyr)
library(tibble)
library(EnhancedVolcano)
library(ggsci)
```

#set colour palette
```{r}
mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)
mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)
#show_col(mycoloursP, labels =F)
```
# load objects
```{r}
mglia <- readRDS(here("data/repeat_with_no_schizophrenia/microglia.RDS"))
astro <-  readRDS(here("data/repeat_with_no_schizophrenia/astrocytes.RDS"))
oligo <-  readRDS(here("data/repeat_with_no_schizophrenia/oligodendrocytes.RDS"))
```
# OPC_Pine
```{r}

opc_pine <- oligo %>% filter(ol_clusters_named_without_skf == "OPC_Pine")
DefaultAssay(opc_pine) <- "RNA"
```


```{r}
Idents(opc_pine) <- "donor_status"

tissues<- unique(opc_pine$Tissue)
de_big <- data.frame()
for(tissue in tissues){
  print(tissue)
  opc_pine_subset <- opc_pine %>% filter(Tissue == tissue)
  de <- FindMarkers(opc_pine_subset, ident.1 = "HD", logfc.threshold = 0.8, test.use = "MAST"  )
  de <- de %>% mutate(tissue = tissue) %>% 
    filter(p_val_adj < 0.05) %>% 
    rownames_to_column("gene")
  write.csv(de, here("outs", "repeat_with_no_schizophrenia", "chaperons", paste0( "HDvsCntrl_DE_in_OPC_Pine_", tissue,".csv")))

  de_big <- rbind(de_big, de)
}
write.csv(de_big, here("outs", "repeat_with_no_schizophrenia", "chaperons", "OPC_Pine_HDvsCntrl_DE_in_each_tissue.csv"))
```

# astro_Thyme

```{r}
astro$as_clusters_named <- Idents(astro)
astro_thyme <- astro %>% filter(as_clusters_named_without_skf == "Astro_Thyme")

```


```{r}
Idents(astro_thyme) <- "donor_status"

tissues<- unique(astro_thyme$Tissue)
de_big <- data.frame()
for(tissue in tissues){
  print(tissue)
  astro_thyme_subset <- astro_thyme %>% filter(Tissue == tissue)
  de <- FindMarkers(astro_thyme_subset, ident.1 = "HD", logfc.threshold = 0.8, test.use = "MAST"  )
  de <- de %>% mutate(tissue = tissue) %>% 
    filter(p_val_adj < 0.05) %>% 
    rownames_to_column("gene")
  write.csv(de, here("outs", "repeat_with_no_schizophrenia", "chaperons", paste0( "HDvsCntrl_DE_in_astro_thyme_", tissue,".csv")))

  de_big <- rbind(de_big, de)
}
write.csv(de_big, here("outs", "repeat_with_no_schizophrenia", "chaperons", "astro_thyme_HDvsCntrl_DE_in_each_tissue.csv"))
```

# mglia_violet
```{r}

mglia_violet <- mglia %>% filter(as_clusters_named_without_skf == "Mglia_Violet")
```

```{r}
Idents(mglia_violet) <- "donor_status"

tissues<- unique(mglia_violet$Tissue)
de_big <- data.frame()
for(tissue in tissues){
  print(tissue)
  mglia_violet_subset <- mglia_violet %>% filter(Tissue == tissue)
  de <- FindMarkers(mglia_violet_subset, ident.1 = "HD", logfc.threshold = 0.8, test.use = "MAST"  )
  de <- de %>% mutate(tissue = tissue) %>% 
    filter(p_val_adj < 0.05) %>% 
    rownames_to_column("gene")
  write.csv(de, here("outs", "repeat_with_no_schizophrenia", "chaperons", paste0( "HDvsCntrl_DE_in_mglia_violet_", tissue,".csv")))

  de_big <- rbind(de_big, de)
}
write.csv(de_big, here("outs", "repeat_with_no_schizophrenia", "chaperons", "mglia_violet_HDvsCntrl_DE_in_each_tissue.csv"))
```

# Explore
Reloading the csvs with shorter names
```{r}
library(readr)
astro_thyme_de <- read_csv(here("outs", "repeat_with_no_schizophrenia", "chaperons", "astro_thyme_HDvsCntrl_DE_in_each_tissue.csv"))
mglia_violet_de <- read_csv(here("outs", "repeat_with_no_schizophrenia", "chaperons", "mglia_violet_HDvsCntrl_DE_in_each_tissue.csv"))
OPC_pine_de <- read_csv(here("outs", "repeat_with_no_schizophrenia", "chaperons", "OPC_Pine_HDvsCntrl_DE_in_each_tissue.csv"))

```

```{r}
mglia_violet_de %>% 
  count(tissue)

mglia_violet_de %>% 
  filter(avg_log2FC > 0) %>% 
  count(tissue)

# this does not make sense if I'm not filtering for + too (just testing)
mglia_violet_de %>% 
  filter(pct.1 > 0.25,
         pct.2 < 0.6) %>% 
  count(tissue)

```
```{r}
astro_thyme_de %>% 
  count(tissue)

astro_thyme_de %>% 
  filter(avg_log2FC > 0) %>% 
  count(tissue)

astro_thyme_de %>% 
  filter(avg_log2FC > 0) %>% 
  filter(tissue == "HC") 
```
```{r}
OPC_pine_de %>% count(tissue)
```

<!-- # common DE -->
<!-- ```{r} -->
<!-- opc_de <- read.csv(here("outs/repeat_with_no_schizophrenia/chaperons/OPC_Pine_HDvsCntrl_DE_in_each_tissue.csv")) -->
<!-- astro_de <- read.csv(here("outs/repeat_with_no_schizophrenia/chaperons/astro_thyme_HDvsCntrl_DE_in_each_tissue.csv")) -->
<!-- mglia_de <- read.csv(here("outs/repeat_with_no_schizophrenia/chaperons/mglia_violet_HDvsCntrl_DE_in_each_tissue.csv")) -->

<!-- intersect(intersect(mglia_de$gene, opc_de$gene), astro_de$gene) -->
<!-- ``` -->
# common cluster markers
```{r}
astro_markers <- read.csv("outs/repeat_with_no_schizophrenia/subset_celltypes/astro_markers_filter.csv")
thyme_markers <- astro_markers %>% filter(cluster == "Astro_Thyme")

mglia_markers <- read.csv("outs/repeat_with_no_schizophrenia/subset_celltypes/mglia_markers_filter_sortbyFC.csv")
violet_markers <- mglia_markers %>% filter(cluster == "Mglia_Violet")
opc_markers <- read.csv("outs/repeat_with_no_schizophrenia/subset_celltypes/oligo_markers_filter_sortbyFC.csv")
pine_markers <- opc_markers %>% filter(cluster == "OPC_Pine")

# compare
# all
length(intersect(intersect(violet_markers$gene, thyme_markers$gene), pine_markers$gene))
# pairwise
length(intersect(violet_markers$gene, thyme_markers$gene))
length(intersect(pine_markers$gene, thyme_markers$gene))
length(intersect(pine_markers$gene, violet_markers$gene))
# how many genes to start with?
dim(violet_markers)
dim(thyme_markers)
dim(pine_markers)

# list of genes any overlapp
overlaps <- unique(c(intersect(violet_markers$gene, thyme_markers$gene),
              intersect(pine_markers$gene, thyme_markers$gene),
               intersect(pine_markers$gene, violet_markers$gene)))
write.csv(overlaps, here("outs/repeat_with_no_schizophrenia/chaperons/pairwise_overlaps.csv"))
```
Check if they are related to HSP (by looking at the GO biological processes database)
 "BAG3"  - yes
"SERPINH1" -yes
"HSPB1"  -  yes
"SLC7A5" -no
"OSBPL3"  - no
"CDC14A" -no  
"TTTY14"   no
"FNBP1L"   no
"NXN"    no
"GPCPD1" - no


```{r}
chaperons <- c("HSPA1B", "HSPA1A", "BAG3", 
               "HSPA9", "HSPA4", "ST13", 
               "HSPA4L", "HSPH1", "CHORDC1", "PTGES3", "PPID", "HSPB1", "DNAJB1", "DNAJB6","DNAJA1","HSPD1","CCT3","CCT4")

DefaultAssay(mglia) <- "RNA"
DotPlot(mglia, chaperons)

DefaultAssay(astro) <- "RNA"
DotPlot(astro, chaperons)

DefaultAssay(oligo) <- "RNA"
DotPlot(oligo, chaperons)
```


# Volcano Plots
<!-- ##mglia -->
<!-- ```{r} -->



<!-- for(i in 1:length(mglia_violet_de$mm_clu)){ -->
<!--   mglia_violet_de$Region[i] <- strsplit(mglia_violet_de$mm_clu, "_")[[i]][1] -->
<!--   mglia_violet_de$Cluster[i] <- strsplit(mglia_violet_de$mm_clu, "_")[[i]][3] -->
<!-- } -->


<!-- keyvals.col <- ifelse( -->
<!--   mglia_violet_de$Region == "HC", mycoloursP[16],  mycoloursP[18]) -->

<!-- names(keyvals.col)[keyvals.col == mycoloursP[16]]<-"HC" -->
<!-- names(keyvals.col)[keyvals.col == mycoloursP[18]]<-"CN" -->


<!-- keyvals.shape <- ifelse( -->
<!--   mglia_violet_de$Region == "CN", 17,16) -->


<!-- names(keyvals.shape)[keyvals.shape == 17] <- 'CN' -->
<!-- names(keyvals.shape)[keyvals.shape == 16] <- 'HC' -->


<!-- int_genes <- c("FKBP4","ST13","AHSA1","HSPE1","HSPB1","DNAJB1","DNAJB6","HSPA4L","PPID","HSPA6","HSPH1","CHORDC1","BAG3","HSP90AB1","HSPA1B","DNAJA1","HSPD1","CCT4","DNAJA4","PTGES3","HSPA4","SERPINH1","IER5","SUGT1", "NEDD4L") -->



<!-- p1 <- EnhancedVolcano(mglia_violet_de, -->
<!--                 lab = mglia_violet_de$gene, -->
<!--                 x = "avg_log2FC", -->
<!--                 y = "p_val_adj",  -->
<!--                 labSize = 3.5, -->
<!--                 FCcutoff = 0.8, -->
<!--                 pCutoff = 0.05, -->
<!--                 pointSize = 3, -->
<!--                 raster = F, -->
<!--                 #shapeCustom = keyvals.shape, -->
<!--                 colCustom = keyvals.col, -->
<!--                 selectLab = unique(int_genes), -->
<!--                 boxedLabels = T,          -->
<!--                 drawConnectors = T, -->
<!--                 #maxoverlapsConnectors = 3, -->
<!--                 labFace = "bold", -->
<!--                 title = NULL, -->
<!--                 subtitle =NULL, -->
<!--                 legendLabSize = 12, -->
<!--                 legendIconSize = 4, -->
<!--                 legendPosition = "right", -->
<!--                 xlim = c((min(mglia_violet_de$avg_log2FC) - .05), (max(mglia_violet_de$avg_log2FC + 0.5))))  -->
<!-- ``` -->




the genes showed are from https://www.nature.com/articles/s41467-021-22369-9#MOESM1
```{r}

chaperon_long_list <- read.csv(here("data/chaperon_list_moesm1.csv"))$Name
```

##microglia
```{r}


violet_de_2tissues <- mglia_violet_de %>% filter(tissue %in% c("HC", "CN"))


keyvals.col <- ifelse(
  violet_de_2tissues$tissue == "HC", mycoloursP[16],  mycoloursP[18])

names(keyvals.col)[keyvals.col == mycoloursP[16]]<-"HC"
names(keyvals.col)[keyvals.col == mycoloursP[18]]<-"CN"


#keyvals.shape <- ifelse(
 # violet_de_2tissues$Region == "CN", 17,16)


#names(keyvals.shape)[keyvals.shape == 17] <- 'CN'
#names(keyvals.shape)[keyvals.shape == 16] <- 'HC'


int_genes <- chaperon_long_list



p1 <- EnhancedVolcano(violet_de_2tissues,
                lab = violet_de_2tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 3.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                max.overlaps = Inf,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                #maxoverlapsConnectors = 3,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(violet_de_2tissues$avg_log2FC) - .05), (max(violet_de_2tissues$avg_log2FC + 0.5)))) 
#ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_mglia_violet_HC_CN_test.pdf"), p1, width = 15, height = 7)
ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_mglia_violet_HC_CN.pdf"), p1, width = 10, height = 5)
```


```{r}


violet_de_3tissues <- mglia_violet_de %>% filter(tissue %in% c("FrCx", "CN", "HC"))


keyvals.col <- case_when(
  violet_de_3tissues$tissue == "HC" ~ mycoloursP[16],  
  violet_de_3tissues$tissue == "FrCx" ~ mycoloursP[8],
  violet_de_3tissues$tissue == "CN"    ~   mycoloursP[18])

names(keyvals.col)[keyvals.col == mycoloursP[16]]<-"HC"
names(keyvals.col)[keyvals.col == mycoloursP[18]]<-"CN"
names(keyvals.col)[keyvals.col == mycoloursP[8]]<-"FrCx"

#keyvals.shape <- ifelse(
 # violet_de_2tissues$Region == "CN", 17,16)


# names(keyvals.shape)[keyvals.shape == 17] <- 'CN'
# names(keyvals.shape)[keyvals.shape == 16] <- 'FrCx'


int_genes <- chaperon_long_list



p1 <- EnhancedVolcano(violet_de_3tissues,
                lab = violet_de_3tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 3.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                colGradient = 
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                #maxoverlapsConnectors = 3,
                max.overlaps = Inf,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(violet_de_3tissues$avg_log2FC) - .05), (max(violet_de_3tissues$avg_log2FC + 0.5)))) 
ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_mglia_violet_HC_CN_FrCx.pdf"), p1, width = 10, height = 5)

# smalller
p2 <- EnhancedVolcano(violet_de_3tissues,
                lab = violet_de_3tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 2.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                #maxoverlapsConnectors = 3,
                max.overlaps = Inf,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(violet_de_3tissues$avg_log2FC) - .05), (max(violet_de_3tissues$avg_log2FC + 0.5)))) 
ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_mglia_violet_HC_CN_FrCx_smaller_labs.pdf"), p2, width = 10, height = 5)
```


```{r}


violet_de_4tissues <- mglia_violet_de 

keyvals.col <- case_when(
  violet_de_4tissues$tissue == "HC" ~ mycoloursP[16],  
  violet_de_4tissues$tissue == "FrCx" ~ mycoloursP[8],
  violet_de_4tissues$tissue == "CN" ~ mycoloursP[18],
  violet_de_4tissues$tissue == "CB" ~ mycoloursP[22])

names(keyvals.col)[keyvals.col == mycoloursP[16]]<-"HC"
names(keyvals.col)[keyvals.col == mycoloursP[18]]<-"CN"
names(keyvals.col)[keyvals.col == mycoloursP[8]]<-"FrCx"
names(keyvals.col)[keyvals.col == mycoloursP[22]]<-"CB"

#keyvals.shape <- ifelse(
 # violet_de_2tissues$Region == "CN", 17,16)


# names(keyvals.shape)[keyvals.shape == 17] <- 'CN'
# names(keyvals.shape)[keyvals.shape == 16] <- 'FrCx'


int_genes <- chaperon_long_list



p1 <- EnhancedVolcano(violet_de_4tissues,
                lab = violet_de_4tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 3.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                #maxoverlapsConnectors = 3,
                max.overlaps = Inf,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(violet_de_4tissues$avg_log2FC) - .05), (max(violet_de_4tissues$avg_log2FC + 0.5)))) 
ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_mglia_violet_HC_CN_FrCx_CB.pdf"), p1, width = 10, height = 5)

# smalller
p2 <- EnhancedVolcano(violet_de_4tissues,
                lab = violet_de_4tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 2.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                #maxoverlapsConnectors = 3,
                max.overlaps = Inf,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(violet_de_4tissues$avg_log2FC) - .05), (max(violet_de_4tissues$avg_log2FC + 0.5)))) 
ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_mglia_violet_HC_CN_CB_FrCx_smaller_labs.pdf"), p2, width = 10, height = 5)

p3 <- EnhancedVolcano(violet_de_4tissues,
                lab = violet_de_4tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 2.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = F, 
                drawConnectors = T,
                #maxoverlapsConnectors = 3,
                max.overlaps = Inf,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                axisLabSize = 18,
                legendLabSize = 18,
                legendIconSize = 6,
                legendPosition = "right",
                xlim = c((min(violet_de_4tissues$avg_log2FC) - .05), (max(violet_de_4tissues$avg_log2FC + 0.5)))) 
ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_mglia_violet_HC_CN_CB_FrCx_no_box.pdf"), p3, width = 10, height = 5)
```
## astro
```{r}


thyme_de_2tissues <- astro_thyme_de %>% filter(tissue %in% c("HC", "FrCx"))


keyvals.col <- ifelse(
  thyme_de_2tissues$tissue == "HC", mycoloursP[16],  mycoloursP[8])

names(keyvals.col)[keyvals.col == mycoloursP[16]]<-"HC"
names(keyvals.col)[keyvals.col == mycoloursP[8]]<-"FrCx"




int_genes <- chaperon_long_list



p1 <- EnhancedVolcano(thyme_de_2tissues,
                lab = thyme_de_2tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 3.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                max.overlaps = Inf,
                #maxoverlapsConnectors = 3,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(thyme_de_2tissues$avg_log2FC) - .05), (max(thyme_de_2tissues$avg_log2FC + 0.5)))) 
ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_astro_thyme_HC_FrCx.pdf"), p1, width = 10, height = 5)
```


```{r}


thyme_de_3tissues <- astro_thyme_de %>% filter(tissue %in% c("FrCx", "CN", "HC"))


keyvals.col <- case_when(
  thyme_de_3tissues$tissue == "HC" ~ mycoloursP[16],  
  thyme_de_3tissues$tissue == "FrCx" ~ mycoloursP[8],
  thyme_de_3tissues$tissue == "CN"    ~   mycoloursP[18])

names(keyvals.col)[keyvals.col == mycoloursP[16]]<-"HC"
names(keyvals.col)[keyvals.col == mycoloursP[18]]<-"CN"
names(keyvals.col)[keyvals.col == mycoloursP[8]]<-"FrCx"

#keyvals.shape <- ifelse(
 # violet_de_2tissues$Region == "CN", 17,16)


# names(keyvals.shape)[keyvals.shape == 17] <- 'CN'
# names(keyvals.shape)[keyvals.shape == 16] <- 'FrCx'


int_genes <- chaperon_long_list



p1 <- EnhancedVolcano(thyme_de_3tissues,
                lab = thyme_de_3tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 3.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                max.overlaps = Inf,
                #maxoverlapsConnectors = 3,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(violet_de_3tissues$avg_log2FC) - .05), (max(thyme_de_3tissues$avg_log2FC + 0.5)))) 


ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_astro_thyme_HC_CN_FrCx.pdf"), p1, width = 10, height = 4)


p1 <- EnhancedVolcano(thyme_de_3tissues,
                lab = thyme_de_3tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 2,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                max.overlaps = Inf,
                #maxoverlapsConnectors = 3,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(violet_de_3tissues$avg_log2FC) - .05), (max(thyme_de_3tissues$avg_log2FC + 0.5)))) 


ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_astro_thyme_HC_CN_FrCx_smaller_labs.pdf"), p1, width = 10, height = 4)
```

## oligos
```{r}


pine_de_3tissues <- OPC_pine_de %>% filter(tissue %in% c("CB", "CN", "HC"))


keyvals.col <- case_when(
  pine_de_3tissues$tissue == "HC" ~ mycoloursP[16],  
  pine_de_3tissues$tissue == "CB" ~ mycoloursP[22],
  pine_de_3tissues$tissue == "CN"    ~   mycoloursP[18])

names(keyvals.col)[keyvals.col == mycoloursP[16]]<-"HC"
names(keyvals.col)[keyvals.col == mycoloursP[18]]<-"CN"
names(keyvals.col)[keyvals.col == mycoloursP[22]]<-"CB"
names(keyvals.col)[keyvals.col == mycoloursP[8]]<-"FrCx"

#keyvals.shape <- ifelse(
 # violet_de_2tissues$Region == "CN", 17,16)


# names(keyvals.shape)[keyvals.shape == 17] <- 'CN'
# names(keyvals.shape)[keyvals.shape == 16] <- 'FrCx'


int_genes <- chaperon_long_list



p1 <- EnhancedVolcano(pine_de_3tissues,
                lab = pine_de_3tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 3.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                max.overlaps = Inf,
                #maxoverlapsConnectors = 3,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(pine_de_3tissues$avg_log2FC) - .05), (max(pine_de_3tissues$avg_log2FC + 0.5)))) 


ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_opc_pine_HC_CN_CB.pdf"), p1, width = 10, height = 4)


p1 <- EnhancedVolcano(pine_de_3tissues,
                lab = pine_de_3tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 2,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                max.overlaps = Inf,
                #maxoverlapsConnectors = 3,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(pine_de_3tissues$avg_log2FC) - .05), (max(pine_de_3tissues$avg_log2FC + 0.5)))) 


ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_opc_pine_HC_CN_CB_smaller_labs.pdf"), p1, width = 10, height = 4)
```
HC 16, CB 22


```{r}


pine_de_4tissues <- OPC_pine_de 


keyvals.col <- case_when(
  pine_de_4tissues$tissue == "HC" ~ mycoloursP[16],  
  pine_de_4tissues$tissue == "CB" ~ mycoloursP[22],
  pine_de_4tissues$tissue == "CN"    ~   mycoloursP[18],
  pine_de_4tissues$tissue == "FrCx" ~ mycoloursP[8])

names(keyvals.col)[keyvals.col == mycoloursP[16]]<-"HC"
names(keyvals.col)[keyvals.col == mycoloursP[18]]<-"CN"
names(keyvals.col)[keyvals.col == mycoloursP[22]]<-"CB"
names(keyvals.col)[keyvals.col == mycoloursP[8]]<-"FrCx"

#keyvals.shape <- ifelse(
 # violet_de_2tissues$Region == "CN", 17,16)


# names(keyvals.shape)[keyvals.shape == 17] <- 'CN'
# names(keyvals.shape)[keyvals.shape == 16] <- 'FrCx'


int_genes <- chaperon_long_list



p1 <- EnhancedVolcano(pine_de_4tissues,
                lab = pine_de_4tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 3.5,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                max.overlaps = Inf,
                #maxoverlapsConnectors = 3,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(pine_de_4tissues$avg_log2FC) - .05), (max(pine_de_4tissues$avg_log2FC + 0.5)))) 


ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_opc_pine_HC_CN_CB_FrCx.pdf"), p1, width = 10, height = 4)


p1 <- EnhancedVolcano(pine_de_4tissues,
                lab = pine_de_4tissues$gene,
                x = "avg_log2FC",
                y = "p_val_adj", 
                labSize = 2,
                FCcutoff = 0.8,
                pCutoff = 0.05,
                pointSize = 3,
                raster = F,
                #shapeCustom = keyvals.shape,
                colCustom = keyvals.col,
                selectLab = unique(int_genes),
                boxedLabels = T,         
                drawConnectors = T,
                max.overlaps = Inf,
                #maxoverlapsConnectors = 3,
                labFace = "bold",
                title = NULL,
                subtitle =NULL,
                legendLabSize = 18,
                legendIconSize = 4,
                legendPosition = "right",
                xlim = c((min(pine_de_4tissues$avg_log2FC) - .05), (max(pine_de_4tissues$avg_log2FC + 0.5)))) 


ggsave(here("outs/repeat_with_no_schizophrenia/chaperons/volcano_plot_chaperons_opc_pine_HC_CN_CB_FrCx_smaller_labs.pdf"), p1, width = 10, height = 4)
```

# Summaries
```{r}
filter(OPC_pine_de, gene %in% int_genes) %>% count(tissue)
filter(mglia_violet_de, gene %in% int_genes) %>% count(tissue)
filter(astro_thyme_de, gene %in% int_genes) %>% count(tissue)
table(OPC_pine_de$tissue)
table(mglia_violet_de$tissue)
table(astro_thyme_de$tissue)

```

